Density Matrix Susceptibility Tensors

### 2018-06-01 papers/1_dmatpol/paper-r1.tex
Revisions in the table and appendix A

Performance descriptors of the \emph{ab initio} and generalized 
models of the density matrix susceptibility are shown in Table~\ref{t:results}. 
%
\begin{table*}[]
\centering
\caption{My caption}
\label{t:results}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
\textbf{}                                                                           & \multicolumn{1}{l|}{} & \textbf{An Initio DMS} & \multicolumn{4}{c|}{\textbf{Generalized DMS}}                 & \textbf{Multipole} \\ \hline
\textbf{}                                                                           & \textbf{Set}          & \textbf{1, 0}          & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W                     & 0.067 (99.80)          & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8} 
                                                                                    & M                     & 1.07 (96.96)           & 1.24 (95.96)  & 0.86 (98.06)  & 1.20 (96.17)  & 0.81 (98.27)  & 0.07 (99.99)       \\ \cline{2-8} 
                                                                                    & S                     & 4.22 (88.87)           & 5.15 (83.44)  & 3.60 (91.92)  & 5.02 (84.25)  & 3.41 (92.73)  & 0.33 (99.93)       \\ \hline
\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W                     & 3.36 (98.70)           & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8} 
                                                                                    & M                     & 14.72 (98.50)          & 7.72 (99.59)  & 6.86 (99.67)  & 6.15 (99.74)  & 5.38 (99.80)  & 14.72 (98.50)      \\ \cline{2-8} 
                                                                                    & S                     & 44.80 (96.92)          & 33.06 (98.32) & 29.49 (98.67) & 28.74 (98.73) & 26.25 (98.94) & 44.80 (96.92)      \\ \hline
\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W                     & 6.22 (58.72)           & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8} 
                                                                                    & M                     & 28.22 (53.90)          & 10.45 (93.68) & 8.83 (95.49)  & 8.25 (96.06)  & 6.57 (97.50)  & 23.94 (66.83)      \\ \cline{2-8} 
                                                                                    & S                     & 78.07 (40.53)          & 43.33 (81.68) & 35.73 (87.55) & 38.80 (85.31) & 31.59 (90.26) & 64.33 (59.62)      \\ \hline
\end{tabular}
\end{table*}

%We first test the accuracy of the density matrix susceptibility tensors that were derived
%in Eq.~\eqref{e:susceptibility-B} from first principles. 
%For this purpose we have considered a water molecule 
%under weak uniform electric fields (Figure~XXX). 
%3D difference one\hyp{}electron density isosurface plots for sample electric fields
%with respect to the unperturbed density (${\bf F} = {\bf 0}$)
%are shown in Figure~XXX. The statistical evaluation of errors for a set of
%100 random electric fields is shown in Figure~XXX. The shape of the basins 
%of decreased/increased density are comparable with the exact result
%obtained from HF/6-311++G** calculations. 
%It can be seen that the theory
%describes the polarization of the electron density distribution qualitatively well:
%However, the accuracy in the density matrix elements is on overall very low which 
%manifests with drammatic errors in the electronic total energy of the system.
%It is due to the approximations used to derive Eq.~\eqref{e:susceptibility-B},
%which neglects quadrupole and higher\hyp{}order induced multipole moments.

@@

Let us start with noticing that
%
\begin{equation}
 \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
\end{equation}
%
Further we have
%
\begin{equation}
 \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_0 \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
\end{equation}
%
But ${\bf C}^\dagger {\bf D}={\bf C}^\dagger {\bf C} {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
AO basis. This implies that
%
\begin{equation}
   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_0) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_0) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_0)
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0})   - \left[ {\bf K} \right]_{i\alpha} ({\bf 0})
 = \left[ {\bf L} \right]_{i\alpha} ({\bf 0}) \;.
\end{equation}
%
Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
are origin independent.
Thus, one can compute dipole integrals with respect to any origin and resulting
susceptibilities ${\bf B}_{\alpha\beta}^{(i)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.

>>>>

Performance descriptors of the \emph{ab initio} and generalized 
models of the density matrix susceptibility are shown in Table~\ref{t:results}. 
\begin{table*}[]
\centering
\caption{My caption}
\label{t:results}
\begin{ruledtabular}
\resizebox{\textwidth}{!}{%
\begin{tabular}{cddcddddddddcdd}
%\hline\hline
\toprule
\multicolumn{1}{l}{} & \multicolumn{2}{c}{\textbf{An Initio DMS}} &  & \multicolumn{8}{c}{\textbf{Generalized DMS}}                                                                                                  &  & \multicolumn{2}{c}{\textbf{Distributed}} \\
\cline{2-3}
\cline{5-12}
%\cline{14-15}
\textbf{Set}         & \multicolumn{2}{c}{\textbf{1, 0}}          &  & \multicolumn{2}{c}{\textbf{1, 0}} & \multicolumn{2}{c}{\textbf{2, 0}} & \multicolumn{2}{c}{\textbf{1, 1}} & \multicolumn{2}{c}{\textbf{2, 1}} &  & \multicolumn{2}{c}{\textbf{Polarizability}}         \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSE (kcal/mol)}}                                                                                                                                                                                             \\ \hline
W                    & 0.067               & (99.80)              &  & 0.076          & (99.74)          & 0.053          & (99.88)          & 0.074          & (99.76)          & 0.050          & (99.89)          &  & 0.004            & (100.00)            \\
M                    & 1.07                & (96.96)              &  & 1.24           & (95.96)          & 0.86           & (98.06)          & 1.20           & (96.17)          & 0.81           & (98.27)          &  & 0.07             & (99.99)             \\
S                    & 4.22                & (88.87)              &  & 5.15           & (83.44)          & 3.60           & (91.92)          & 5.02           & (84.25)          & 3.41           & (92.73)          &  & 0.33             & (99.93)             \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSD (mD)}}                                                                                                                                                                                                   \\ \hline
W                    & 3.36                & (98.70)              &  & 1.03           & (99.88)          & 0.98           & (99.89)          & 0.52           & (99.97)          & 0.47           & (99.98)          &  & 3.36             & (98.70)             \\
M                    & 14.72               & (98.50)              &  & 7.72           & (99.59)          & 6.86           & (99.67)          & 6.15           & (99.74)          & 5.38           & (99.80)          &  & 14.72            & (98.50)             \\
S                    & 44.80               & (96.92)              &  & 33.06          & (98.32)          & 29.49          & (98.67)          & 28.74          & (98.73)          & 26.25          & (98.94)          &  & 44.80            & (96.92)             \\ \hline
\multicolumn{1}{l}{} & \multicolumn{14}{c}{\textbf{RMSQ (D\AA $\times 10^{-3}$)}}                                                                                                                                                                                               \\ \hline
W                    & 6.22                & (58.72)              &  & 1.59           & (97.32)          & 1.49           & (97.62)          & 0.58           & (99.64)          & 0.48           & (99.75)          &  & 5.45             & (68.30)             \\
M                    & 28.22               & (53.90)              &  & 10.45          & (93.68)          & 8.83           & (95.49)          & 8.25           & (96.06)          & 6.57           & (97.50)          &  & 23.94            & (66.83)             \\
S                    & 78.07               & (40.53)              &  & 43.33          & (81.68)          & 35.73          & (87.55)          & 38.80          & (85.31)          & 31.59          & (90.26)          &  & 64.33            & (59.62) \\
\bottomrule
\end{tabular}%
}
\end{ruledtabular}
\end{table*}
We considered three different types of point charge environment that generate
weak (0.001 -- 0.005 a.u.), 
moderate (0.001 -- 0.002 a.u.) and strong (0.002 -- 0.004 a.u.) 
electric fields around the water molecule (see the histograms of the electric field distribution
per atomic centre in Figure~XXX).
-%
-%for a set of varying point charge environments
-%with weak, moderate and strong electric fields are shown in Table~XXX.
\begin{figure*}
\includegraphics[width=\textwidth]{data/dmatpol/water/figure1/fig-1.eps}
\caption{\label{f:results} {\bf Performance of the distributed density matrix susceptibility tensors in moderate
electric fields.}
Calculations of the electric field\hyp{}induced interaction energies (A-C), dipole moments (D-F) and
quadrupole moments (G-I).}
\end{figure*}
%\begin{table*}[]
%\centering
%\caption{My caption}
%\label{t:results}
%\resizebox{\textwidth}{!}{%
%\begin{tabular}{|l|c|c|c|c|c|c|c|}
%\hline
%\textbf{}                                                                           & \multicolumn{1}{l|}{} & \textbf{An Initio DMS} & \multicolumn{4}{c|}{\textbf{Generalized DMS}}                 & \textbf{Multipole} \\ \hline
%\textbf{}                                                                           & \textbf{Set}          & \textbf{1, 0}          & \textbf{1, 0} & \textbf{2, 0} & \textbf{1, 1} & \textbf{2, 1} & \textbf{-}         \\ \hline
%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSE\\ (kcal/mol)\end{tabular}}} & W                     & 0.067 (99.80)          & 0.076 (99.74) & 0.053 (99.88) & 0.074 (99.76) & 0.050 (99.89) & 0.004 (100.00)     \\ \cline{2-8} 
%                                                                                    & M                     & 1.07 (96.96)           & 1.24 (95.96)  & 0.86 (98.06)  & 1.20 (96.17)  & 0.81 (98.27)  & 0.07 (99.99)       \\ \cline{2-8} 
%                                                                                    & S                     & 4.22 (88.87)           & 5.15 (83.44)  & 3.60 (91.92)  & 5.02 (84.25)  & 3.41 (92.73)  & 0.33 (99.93)       \\ \hline
%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSD\\ (mD)\end{tabular}}}       & W                     & 3.36 (98.70)           & 1.03 (99.88)  & 0.98 (99.89)  & 0.52 (99.97)  & 0.47 (99.98)  & 3.36 (98.70)       \\ \cline{2-8} 
%                                                                                    & M                     & 14.72 (98.50)          & 7.72 (99.59)  & 6.86 (99.67)  & 6.15 (99.74)  & 5.38 (99.80)  & 14.72 (98.50)      \\ \cline{2-8} 
%                                                                                    & S                     & 44.80 (96.92)          & 33.06 (98.32) & 29.49 (98.67) & 28.74 (98.73) & 26.25 (98.94) & 44.80 (96.92)      \\ \hline
%\multirow{3}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}RMSQ\\ (DA10-3)\end{tabular}}}   & W                     & 6.22 (58.72)           & 1.59 (97.32)  & 1.49 (97.62)  & 0.58 (99.64)  & 0.48 (99.75)  & 5.45 (68.30)       \\ \cline{2-8} 
%                                                                                    & M                     & 28.22 (53.90)          & 10.45 (93.68) & 8.83 (95.49)  & 8.25 (96.06)  & 6.57 (97.50)  & 23.94 (66.83)      \\ \cline{2-8} 
%                                                                                    & S                     & 78.07 (40.53)          & 43.33 (81.68) & 35.73 (87.55) & 38.80 (85.31) & 31.59 (90.26) & 64.33 (59.62)      \\ \hline
%\end{tabular}%
%}
%\end{table*}
%We first test the accuracy of the density matrix susceptibility tensors that were derived
%in Eq.~\eqref{e:susceptibility-B} from first principles. 
%For this purpose we have considered a water molecule 
%under weak uniform electric fields (Figure~XXX). 
%3D difference one\hyp{}electron density isosurface plots for sample electric fields
%with respect to the unperturbed density (${\bf F} = {\bf 0}$)
%are shown in Figure~XXX. The statistical evaluation of errors for a set of
%100 random electric fields is shown in Figure~XXX. The shape of the basins 
%of decreased/increased density are comparable with the exact result
%obtained from HF/6-311++G** calculations. 
%It can be seen that the theory
%describes the polarization of the electron density distribution qualitatively well:
%However, the accuracy in the density matrix elements is on overall very low which 
%manifests with drammatic errors in the electronic total energy of the system.
%It is due to the approximations used to derive Eq.~\eqref{e:susceptibility-B},
%which neglects quadrupole and higher\hyp{}order induced multipole moments.

@@

Let us start with noticing that
%
\begin{equation}
 \left[ {\bf M} \right]_{i\alpha} ({\bf r}_Q) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger \right]_{i\alpha}  \;.
\end{equation}
%
Further we have
%
\begin{equation}
 \left[ {\bf K} \right]_{i\alpha} ({\bf r}_Q) 
 = \left[ {\bf K} \right]_{i\alpha} ({\bf 0}) - {\bf r}_Q \left[ {\bf C}^\dagger {\bf D} \right]_{i\alpha} \;.
\end{equation}
%
But ${\bf C}^\dagger {\bf D}={\bf C}^\dagger {\bf C} {\bf C}^\dagger={\bf C}^\dagger$ in orthogonal
AO basis. This implies that
%
\begin{multline}
   \left[ {\bf L} \right]_{i\alpha} ({\bf r}_Q) 
 = \left[ {\bf M} \right]_{i\alpha} ({\bf r}_Q) - \left[ {\bf K} \right]_{i\alpha} ({\bf r}_Q) \\
 = \left[ {\bf M} \right]_{i\alpha} ({\bf 0})   - \left[ {\bf K} \right]_{i\alpha} ({\bf 0})
 = \left[ {\bf L} \right]_{i\alpha} ({\bf 0}) \;.
\end{multline}
%
Therefore, it is proved that the polarization\hyp{}induced distributed dipole moments 
defined in Eq.~\eqref{e:dmu-4-exact.linear-approximation} 
are origin independent.
Thus, one can compute dipole integrals with respect to any origin and resulting
susceptibilities ${\bf B}_{\alpha\beta}^{(i)}$ from Eq.~\eqref{e:susceptibility-B} will be uniquely defined.
